Building Sustainable Software - The CSDMS Approach

### 2014-07-14 paper.tex
Fixed all usages of \csdms{}.

\csdms, The Community Surface Dynamics Modeling System, is an NSF funded
project whose focus is to aid a diverse community of earth and ocean system
model \emph{users} and \emph{developers} to use and create robust software
quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
earth-system models and tools to an international (67 country) community
with the goal of building the set of tools necessary to model the
earth system. \csdms{} is a place modelers look to for access to hundreds open
source surface-dynamics models and tools, as well as model metadata. Such a
model repository increases model transparency and helps to eliminate the
duplication of effors by presenting the current state of modeling efforts.
To increase software sustainability, composability and interoperability \csdms{}
promotes standards that define common modeling interfaces, semantic mediation
between models, and model metadata. Through online resources and workshops,
\csdms{} promotes software engineering best practices, which are foreign to many
developers within our modeling community. For example, version control, unit
testing, continious integration, test-driven development, well-written
\emph{clean code} are all topics of the educational mission of \csdms{}.

@@

\csdms{} has built an online model repository that not only contains model
source code but also documents the submitted code as model metadata, runs
basic smoke-tests of the models, (optionally) maintains the source code in a
\csdms-hosted version control system, and provides easy downloads of the
original source. The repository now contains over 200 models and tools (over
4M lines of code) and has recieved over 13k model downloads.

@@

The \bmi{} identifies the entry points into software components that provides a
calling appication with the necessary level of control to connect components.
\csdms, as well as other modeling frameworks such as
ESMF~\cite{hill2004architecture}, OpenMI~\cite{gregersen2007openmi}, and
OMS~\cite{david2002object}, identifies an interface that, at a minimum,
provides functionality to \emph{initialize}, \emph{update}, and
\emph{finalize} a component. However, for more complicated (and 2-way)
coupling components must implement the \bmi{} methods providing data access,
and model metadata.

@@

Another important goal of CSDMS is to enable the rapid development and application of linked dynamical models tailored to specific landscape-basin evolution problems and at specific temporal and spatial scales. Interdisciplinarity and coupling is a critical step towards advancing predictive modeling~\cite{voinov2010community}.

CSDMS develops a modeling environment containing a suite of integrated, ever-improving software modules, with expanded functionality beyond the original stand-alone codes.

@@

\csdms{} constantly evaluates the effectiveness of its tools and standards
through user feedback from clinics and tutorials and one-on-one interviews
with \csdms{} members.  Modeling courses at the University of Colorado,
clinics at the International Summer Institute at the University of Minnesota,
and clinics at annual CSDMS Annual Meetings are the dominant forums for
this testing and feedback.

>>>>

\csdms{}, The Community Surface Dynamics Modeling System, is an NSF funded
project whose focus is to aid a diverse community of earth and ocean system
model \emph{users} and \emph{developers} to use and create robust software
quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
earth-system models and tools to an international (67 country) community
with the goal of building the set of tools necessary to model the
earth system. \csdms{} is a place modelers look to for access to hundreds open
source surface-dynamics models and tools, as well as model metadata. Such a
model repository increases model transparency and helps to eliminate the
duplication of effors by presenting the current state of modeling efforts.
To increase software sustainability, composability and interoperability \csdms{}
promotes standards that define common modeling interfaces, semantic mediation
between models, and model metadata. Through online resources and workshops,
\csdms{} promotes software engineering best practices, which are foreign to many
developers within our modeling community. For example, version control, unit
testing, continious integration, test-driven development, well-written
\emph{clean code} are all topics of the educational mission of \csdms{}.

@@

\csdms{} has built an online model repository that not only contains model
source code but also documents the submitted code as model metadata, runs
basic smoke-tests of the models, (optionally) maintains the source code in a
\csdms{}-hosted version control system, and provides easy downloads of the
original source. The repository now contains over 200 models and tools (over
4M lines of code) and has recieved over 13k model downloads.

@@

The \bmi{} identifies the entry points into software components that provides a
calling appication with the necessary level of control to connect components.
\csdms{}, as well as other modeling frameworks such as
ESMF~\cite{hill2004architecture}, OpenMI~\cite{gregersen2007openmi}, and
OMS~\cite{david2002object}, identifies an interface that, at a minimum,
provides functionality to \emph{initialize}, \emph{update}, and
\emph{finalize} a component. However, for more complicated (and 2-way)
coupling components must implement the \bmi{} methods providing data access,
and model metadata.

@@

Another important goal of \csdms{} is to enable the rapid development and application of linked dynamical models tailored to specific landscape-basin evolution problems and at specific temporal and spatial scales. Interdisciplinarity and coupling is a critical step towards advancing predictive modeling~\cite{voinov2010community}.

\csdms{} develops a modeling environment containing a suite of integrated, ever-improving software modules, with expanded functionality beyond the original stand-alone codes.

@@

\csdms{} constantly evaluates the effectiveness of its tools and standards
through user feedback from clinics and tutorials and one-on-one interviews
with \csdms{} members.  Modeling courses at the University of Colorado,
clinics at the International Summer Institute at the University of Minnesota,
and clinics at annual \csdms{} Annual Meetings are the dominant forums for
this testing and feedback.
